### 2015-12-17 doc/thesis/intro.tex
Reading and writing, plus cross-validation for ordinary lm

In evolutionary biology, a \defn{phylogeny}, or \defn{phylogenetic tree}, is a
graphical representation of the evolutionary relationships among a group of
organisms or species. Phylogenies consist of nodes, representing organisms or
species (generally, \defn{taxa}), and edges or branches, which link taxa to
their ancestors and descendants. The \defn{tips} of the phylogeny, that is, the
nodes without any descendants, correspond to \defn{extant}, or observed, taxa.
The \defn{internal nodes} correspond to their extinct common ancestors. The
phylogeny may have a \defn{root}, which is a node with no descendants
distinguished as the most recent common ancestor of all the extant taxa. When
such a root exists, the tree is referred to as being \defn{rooted}; otherwise,
it is \defn{unrooted}. The structural arrangement of nodes and edges in the
tree is referred to as its \defn{topology}. The branches of the phylogeny may
have lengths, which can represent either the genetic distance, or the calendar
time, between ancestors and their descendants. Phylogenies whose branch lengths
are in units of calendar time are referred to as \defn{time-scaled}. In the
case of a rooted tree with branch lengths in calendar time, the internal nodes
can be mapped onto a timeline by using the tips of the phylogeny, which usually
map to the present day, as a reference point. These \defn{branching times}
correspond to the times at which speciation events
occurred~\autocite{nee1992tempo}. These concepts are illustrated in
Figure~\ref{fig:phylogeny}.

\begin{figure}
  \label{fig:phylogeny}
  \includegraphics{phylogeny}
  \caption{Cartoon illustration of a rooted, time-scaled phylogeny relating
  several domestic animals. The tips of the phylogeny, which represent extant
  taxa, are placed at the present day on the time axis. Internal nodes,
  representing extinct common ancestors to the extant taxa, fall in the past,
  at branching times corresponding to speciation events. The topology of the
  tree indicates that cats and dogs are the most closely related pair of
  species, whereas fish is most distantly related to any other node in the
  tree.}
\end{figure}

@@

Now, consider a phylogeny not of macro-organisms, but of RNA viruses. Many of
these, such as the human immuno-deficiency virus (HIV), evolve rapidy enough
that populations can accumulate measurable genetic changes on time scales of
months or years~\autocite{drummond2003measurably}. Because viruses cannot exist
for long outside their hosts, the host replaces the outside world as the
``environment'' in which evolution takes place. In this light, striking
parallels can be drawn between immunological and epidemiological processes at
the host level, and the evolutionary processes studied in population
genetics~\autocite{grenfell2004unifying}. For example, both the host immune
response~\autocite{??} and the presence of anti-retroviral drugs~\autocite{??}
exert directional selection on the viral population within a single host.
Transmission, where a sub-population of virions from the donor host is
``geographically'' isolated in the recpient, closely mirrors the allopatric
speciation process and causes a branching point in the viral phylogeny
(although this may not correspond exactly to the time of transmission, as
discussed in the next section). Consequently, any factor which influences
transmission dynamics, such as the virus' mode of transmission or the
population structure of its hosts, in turn affects the shape of the phylogeny.
The study of this interaction between host-level factors and viral phylogies is
known as \defn{phylodynamics}~\autocite{grenfell2004unifying}.

>>>>

\subsubsection{Species trees}

In evolutionary biology, the evolutionary relationships among a group of
organisms or species (generally, \defn{taxa}) are often represented graphically
in the form of a \defn{species tree}. These trees are often called
\defn{phylogenies}, but to avoid ambiguity, we will reserve that term for gene
trees as discussed below. Species trees consist of nodes, representing taxa,
and edges or branches, which connect taxa to their evolutionary ancestors and
descendants. The \defn{tips} of a species tree, that is, the nodes without any
descendants, correspond to \defn{extant}, or observed, taxa, while the
\defn{internal nodes} correspond to their extinct common ancestors. Species
trees may have a \defn{root}, which is a node with no descendants distinguished
as the most recent common ancestor of all the extant taxa. When such a root
exists, the tree is referred to as being \defn{rooted}; otherwise, it is
\defn{unrooted}. The structural arrangement of nodes and edges in the tree is
referred to as its \defn{topology}. The branches of the tree may have
associated lengths, which, represent the amount of calendar time elapsed
between ancestors and their descendants. When these branch lengths are present,
the internal nodes can be mapped onto a timeline by using the tips of the tree,
which usually map to the present day, as a reference point. These
\defn{branching times} correspond to the historical times at which speciation
events occurred~\autocite{nee1992tempo}. These concepts are illustrated in
Figure~\ref{fig:speciestree}.

\begin{figure}[h]
  \centering
  \label{fig:speciestree}
  \includegraphics{speciestree}
  \caption{Illustration of a rooted species tree. The tips of the tree, which
    represent extant taxa, are placed at the present day on the time axis.
    Internal nodes, representing extinct common ancestors to the extant taxa,
    fall in the past, at branching times corresponding to speciation events.
    The topology of the tree indicates that cats and dogs are the most closely
    related pair of species, whereas fish is most distantly related to any
    other node in the tree.}
\end{figure}

@@

Now, consider a phylogeny not of macro-organisms, but of RNA viruses. Many of
these, such as the human immuno-deficiency virus (HIV), evolve rapidy enough
that populations can accumulate measurable genetic changes on time scales of
months or years~\autocite{drummond2003measurably}. Because viruses cannot exist
for long outside their hosts, the host replaces the outside world as the
``environment'' in which evolution takes place. In this light, striking
parallels can be drawn between immunological and epidemiological processes at
the host level, and the evolutionary processes studied in population
genetics~\autocite{grenfell2004unifying}. For example, both the host immune
response~\autocite{mcmichael2001cellular} and the presence of anti-retroviral
drugs~\autocite{??} exert directional selection on the viral population within
a single host. Transmission, where a sub-population of virions from the donor
host is ``geographically'' isolated in the recpient, closely mirrors the
allopatric speciation process and causes a branching point in the viral
phylogeny (although this may not correspond exactly to the time of
transmission, as discussed in the next section). Consequently, any factor which
influences transmission dynamics, such as the virus' mode of transmission or
the population structure of its hosts, in turn affects the shape of the
phylogeny. The study of this interaction between host-level factors and viral
phylogies is known as \defn{phylodynamics}~\autocite{grenfell2004unifying}.

### 2015-12-17 doc/thesis/methods.tex
Reading and writing, plus cross-validation for ordinary lm

\subsubsection{Management of computational experiments with combinatorially
expanded parameters}

This reasearch involved many computational experiments performed with all
possible combinations of several parameters. For example, to identify the best
combination of tree kernel parameters (see \ref{subsubsec:kernel} below), we
considered at least three values each of the parameters $\lambda$ and $\sigma$,
corresponding to at least nine possible combinations of these.

@@

\subsubsection{ABC-SMC}

>>>>

@@

\subsubsection{Approximate Bayesian computation}

\subsection{Applications}

\subsubsection{HIV in British Columubia}
